---
title: "STAT 184 Final Project"
author: "Yiwen Huang"
date: "Sunday by 11:59pm"
output: html_notebook
---


# Set Up
```{r, message = FALSE, warning = FALSE}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(DataComputing)
library(mosaic)
library(tidyverse)
library(lubridate)
library(readxl)
library(rvest)
library(knitr)
library(ggplot2)
library(usmap)
library(dendextend)
```


# Research Question

### Are Trap Laws associated with lower reported legal abortion number?

* Trap Law: Targeted restrictions on abortion providers (TRAP) laws are costly, severe, and medically unnecessary requirements imposed on abortion providers and woman's health centers. Often pushed by anti-abortion politicians under the guise of "woman's health," the real aim of TRAP laws is to shut down abortion providers and make it more difficult for people to access abortion.


# Data Access

### Import Data Source

#### 1. Abortion Surveillance — United States, 2015
```{r, message = FALSE, warning = FALSE}
# Assign the url for source 1 to variable url1
url1 <- "https://www.cdc.gov/mmwr/volumes/67/ss/ss6713a1.htm?fbclid=IwAR1RVZmh6lh4UWS6z6jL_fwbVm0pfCJ9H9azAK8TLS5hwyss-8qMi5uG9vo"

# Scrape all HTML tables found on the web page into a "list" object named Abortion2015 in the R environment
Abortion2015 <- url1 %>%
  read_html(header = TRUE) %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)
```

#### 2. Abortion Surveillance — United States, 2016
```{r, message = FALSE, warning = FALSE}
# Assign the url for source 2 to variable url2
url2 <- "https://www.cdc.gov/mmwr/volumes/68/ss/ss6811a1.htm#T1_down"

# Scrape all HTML tables found on the web page into a "list" object named Abortion2016 in the R environment
Abortion2016 <- url2 %>%
  read_html(header = TRUE) %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)
```

#### 3. Anti_Choice Law by States
```{r, message = FALSE, warning = FALSE}
# Use function read-excel from package readxl to read a local excel file into an R object named TrapLaw
TrapLaw <- read_excel("/Users/yiwen/Desktop/FALL 2020/STAT 184/Final Project/TrapLaw.xlsx")
```


### Inspecting Data Intake

#### 1. Abortion Surveillance — United States, 2015
```{r, message = FALSE, warning = FALSE}
StatesReport2015_Raw <- Abortion2015[[2]]
head(StatesReport2015_Raw)
tail(StatesReport2015_Raw)
glimpse(StatesReport2015_Raw)
str(StatesReport2015_Raw)

AgeGroup2015_Raw <- Abortion2015[[3]]
head(AgeGroup2015_Raw)
tail(AgeGroup2015_Raw)
glimpse(AgeGroup2015_Raw)
str(AgeGroup2015_Raw)
```

#### 2. Abortion Surveillance — United States, 2016
```{r, message = FALSE, warning = FALSE}
AnnualReport_Raw <- Abortion2016[[1]]
head(AnnualReport_Raw)
tail(AnnualReport_Raw)
glimpse(AnnualReport_Raw)
str(AnnualReport_Raw)

StatesReport2016_Raw <- Abortion2016[[2]]
head(StatesReport2016_Raw)
tail(StatesReport2016_Raw)
glimpse(StatesReport2016_Raw)
str(StatesReport2016_Raw)

AgeGroup2016_Raw <- Abortion2016[[3]]
head(AgeGroup2016_Raw)
tail(AgeGroup2016_Raw)
glimpse(AgeGroup2016_Raw)
str(AgeGroup2016_Raw)
```


#### 3. Anti_Choice Law by States
```{r, message = FALSE, warning = FALSE}
head(TrapLaw)
tail(TrapLaw)
glimpse(TrapLaw)
str(TrapLaw)
```


# Data Wrangling

#### 1. StatesReport2015_Clean

**Variables**

* `State/Area`
* `Residence Number`
* `Residence Rate`
* `Residence Ratio`
* `Occurrence Number`
* `Occurrence Rate`
* `Occurrence Ratio`
* `% obtained by out-of-state residents`
```{r, message = FALSE, warning = FALSE}
# Rename all the variables in StatesReport2015_Raw
names(StatesReport2015_Raw)[2] <-paste("Residence Number")
names(StatesReport2015_Raw)[3] <-paste("Residence Rate")
names(StatesReport2015_Raw)[4] <-paste("Residence Ratio")
names(StatesReport2015_Raw)[5] <-paste("Occurrence Number")
names(StatesReport2015_Raw)[6] <-paste("Occurrence Rate")
names(StatesReport2015_Raw)[7] <-paste("Occurrence Ratio")

# Remove the redundant row of column names, create regular expression to grasp all the numerical cell values and convert those values from strings to numbers
StatesReport2015_Clean <-
  StatesReport2015_Raw %>%
  filter(row_number() > 1, row_number() < 55) %>%
  mutate(`Residence Number` = as.numeric(gsub(pattern = ",", replacement = "", x = `Residence Number`)),
         `Residence Rate` = as.numeric(gsub(pattern = ",", replacement = "", x = `Residence Rate`)),
         `Residence Ratio` = as.numeric(gsub(pattern = ",", replacement = "", x = `Residence Ratio`)),
         `Occurrence Number` = as.numeric(gsub(pattern = ",", replacement = "", x = `Occurrence Number`)),
         `Occurrence Rate` = as.numeric(gsub(pattern = ",", replacement = "", x = `Occurrence Rate`)),
         `Occurrence Ratio` = as.numeric(gsub(pattern = ",", replacement = "", x = `Occurrence Ratio`))) %>%
  extract(col = `% obtained by out-of-state residents§`, into = "% obtained by out-of-state residents", regex = "(\\d+\\.\\d+)" ) %>%
  mutate(`% obtained by out-of-state residents` = as.numeric(`% obtained by out-of-state residents`))

str(StatesReport2015_Clean)
```

#### 2. AgeGroup2015_Clean

**Variables**

* `State/Area`
* `Age group yrs (<15)` 
* `Age group yrs (15–19)`
* `Age group yrs (20–24)`
* `Age group yrs (25–29)`
* `Age group yrs (30–34)`
* `Age group yrs (>=40)`
* `Total abortions reported by known age`
```{r, message = FALSE, warning = FALSE}
# Rename all the variables in AgeGroup2015_Raw
names(AgeGroup2015_Raw)[2] <-paste("Age group yrs (<15)")
names(AgeGroup2015_Raw)[3] <-paste("Age group yrs (15–19)")
names(AgeGroup2015_Raw)[4] <-paste("Age group yrs (20–24)")
names(AgeGroup2015_Raw)[5] <-paste("Age group yrs (25–29)")
names(AgeGroup2015_Raw)[6] <-paste("Age group yrs (30–34)")
names(AgeGroup2015_Raw)[7] <-paste("Age group yrs (35–39)")
names(AgeGroup2015_Raw)[8] <-paste("Age group yrs (>=40)")

# Remove the redundant row of column names, create regular expression to grasp all the numerical cell values and convert those values from strings to numbers
AgeGroup2015_Clean <-
  AgeGroup2015_Raw %>%
  filter(row_number() > 2, row_number() < 51) %>%
  extract(col = `Age group yrs (<15)`, into = "Age group yrs (<15)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (<15)` = as.numeric(`Age group yrs (<15)`)) %>%
  mutate(`Age group yrs (15–19)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (15–19)`)) %>%
  extract(col = `Age group yrs (15–19)`, into = "Age group yrs (15–19)", regex ="(^\\d+)") %>%
  mutate(`Age group yrs (15–19)` = as.numeric(`Age group yrs (15–19)`)) %>%
  mutate(`Age group yrs (20–24)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (20–24)`)) %>%
  extract(col = `Age group yrs (20–24)`, into = "Age group yrs (20–24)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (20–24)` = as.numeric(`Age group yrs (20–24)`)) %>%
  mutate(`Age group yrs (25–29)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (25–29)`)) %>%
  extract(col = `Age group yrs (25–29)`, into = "Age group yrs (25–29)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (25–29)` = as.numeric(`Age group yrs (25–29)`)) %>%
  mutate(`Age group yrs (30–34)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (30–34)`)) %>%
  extract(col = `Age group yrs (30–34)`, into = "Age group yrs (30–34)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (30–34)` = as.numeric(`Age group yrs (30–34)`)) %>%
  mutate(`Age group yrs (35–39)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (35–39)`)) %>%
  extract(col = `Age group yrs (35–39)`, into = "Age group yrs (35–39)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (35–39)` = as.numeric(`Age group yrs (35–39)`)) %>%
  mutate(`Age group yrs (>=40)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (>=40)`)) %>%
  extract(col = `Age group yrs (>=40)`, into = "Age group yrs (>=40)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (>=40)` = as.numeric(`Age group yrs (>=40)`)) %>%
  mutate(`Total abortions reported by known age` = gsub(pattern = ",", replacement = "", x = `Total abortions reported by known age`)) %>%
  extract(col = `Total abortions reported by known age`, into = "Total abortions reported by known age", regex ="(^\\d+)") %>%
  mutate(`Total abortions reported by known age` = as.numeric(`Total abortions reported by known age`))

str(AgeGroup2015_Clean)
```

#### 3. StatesReport2016_Clean

**Variables**

* `State/Area`
* `Residence Number`
* `Residence Rate`
* `Residence Ratio`
* `Occurrence Number`
* `Occurrence Rate`
* `Occurrence Ratio`
* `% obtained by out-of-state residents`
```{r, message = FALSE, warning = FALSE}
# Rename all the variables in AgeGroup2016_Raw
names(StatesReport2016_Raw)[2] <-paste("Residence Number")
names(StatesReport2016_Raw)[3] <-paste("Residence Rate")
names(StatesReport2016_Raw)[4] <-paste("Residence Ratio")
names(StatesReport2016_Raw)[5] <-paste("Occurrence Number")
names(StatesReport2016_Raw)[6] <-paste("Occurrence Rate")
names(StatesReport2016_Raw)[7] <-paste("Occurrence Ratio")

# Remove the redundant row of column names, create regular expression to grasp all the numerical cell values and convert those values from strings to numbers
StatesReport2016_Clean <-
  StatesReport2016_Raw %>%
  filter(row_number() > 1, row_number() < 55) %>%
  rename(`% obtained by out-of-state residents` = `% obtained by out-of-state residents§`) %>%
  mutate(`Residence Number` = as.numeric(gsub(pattern = ",", replacement = "", x = `Residence Number`)),
         `Residence Rate` = as.numeric(gsub(pattern = ",", replacement = "", x = `Residence Rate`)),
         `Residence Ratio` = as.numeric(gsub(pattern = ",", replacement = "", x = `Residence Ratio`)),
         `Occurrence Number` = as.numeric(gsub(pattern = ",", replacement = "", x = `Occurrence Number`)),
         `Occurrence Rate` = as.numeric(gsub(pattern = ",", replacement = "", x = `Occurrence Rate`)),
         `Occurrence Ratio` = as.numeric(gsub(pattern = ",", replacement = "", x = `Occurrence Ratio`))) %>%
  extract(col = `% obtained by out-of-state residents`, into = "% obtained by out-of-state residents", regex = "(\\d+\\.\\d+)" ) %>%
  mutate(`% obtained by out-of-state residents` = as.numeric(`% obtained by out-of-state residents`))

str(StatesReport2016_Clean)
```

#### 4. AgeGroup2016_Clean

**Variables**

* `State/Area`
* `Age group yrs (<15)` 
* `Age group yrs (15–19)`
* `Age group yrs (20–24)`
* `Age group yrs (25–29)`
* `Age group yrs (30–34)`
* `Age group yrs (>=40)`
* `Total abortions reported by known age`
```{r, message = FALSE, warning = FALSE}
# Rename all the variables in AgeGroup2016_Raw
names(AgeGroup2016_Raw)[1] <-paste("State/Area")
names(AgeGroup2016_Raw)[2] <-paste("Age group yrs (<15)")
names(AgeGroup2016_Raw)[3] <-paste("Age group yrs (15–19)")
names(AgeGroup2016_Raw)[4] <-paste("Age group yrs (20–24)")
names(AgeGroup2016_Raw)[5] <-paste("Age group yrs (25–29)")
names(AgeGroup2016_Raw)[6] <-paste("Age group yrs (30–34)")
names(AgeGroup2016_Raw)[7] <-paste("Age group yrs (35–39)")
names(AgeGroup2016_Raw)[8] <-paste("Age group yrs (>=40)")
names(AgeGroup2016_Raw)[9] <-paste("Total abortions reported by known age")

# Remove the redundant row of column names, create regular expression to grasp all the numerical cell values and convert those values from strings to numbers
AgeGroup2016_Clean <-
  AgeGroup2016_Raw %>%
  filter(row_number() > 3, row_number() < 51) %>%
  extract(col = `Age group yrs (<15)`, into = "Age group yrs (<15)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (<15)` = as.numeric(`Age group yrs (<15)`)) %>%
  mutate(`Age group yrs (15–19)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (15–19)`)) %>%
  extract(col = `Age group yrs (15–19)`, into = "Age group yrs (15–19)", regex ="(^\\d+)") %>%
  mutate(`Age group yrs (15–19)` = as.numeric(`Age group yrs (15–19)`)) %>%
  mutate(`Age group yrs (20–24)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (20–24)`)) %>%
  extract(col = `Age group yrs (20–24)`, into = "Age group yrs (20–24)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (20–24)` = as.numeric(`Age group yrs (20–24)`)) %>%
  mutate(`Age group yrs (25–29)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (25–29)`)) %>%
  extract(col = `Age group yrs (25–29)`, into = "Age group yrs (25–29)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (25–29)` = as.numeric(`Age group yrs (25–29)`)) %>%
  mutate(`Age group yrs (30–34)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (30–34)`)) %>%
  extract(col = `Age group yrs (30–34)`, into = "Age group yrs (30–34)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (30–34)` = as.numeric(`Age group yrs (30–34)`)) %>%
  mutate(`Age group yrs (35–39)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (35–39)`)) %>%
  extract(col = `Age group yrs (35–39)`, into = "Age group yrs (35–39)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (35–39)` = as.numeric(`Age group yrs (35–39)`)) %>%
  mutate(`Age group yrs (>=40)` = gsub(pattern = ",", replacement = "", x = `Age group yrs (>=40)`)) %>%
  extract(col = `Age group yrs (>=40)`, into = "Age group yrs (>=40)", regex = "(^\\d+)") %>%
  mutate(`Age group yrs (>=40)` = as.numeric(`Age group yrs (>=40)`)) %>%
  mutate(`Total abortions reported by known age` = gsub(pattern = ",", replacement = "", x = `Total abortions reported by known age`)) %>%
  extract(col = `Total abortions reported by known age`, into = "Total abortions reported by known age", regex ="(^\\d+)") %>%
  mutate(`Total abortions reported by known age` = as.numeric(`Total abortions reported by known age`))
 
str(AgeGroup2016_Clean) 
```

#### 5. AnnualReport_Clean

**Variables**

* `Year`
* `Selected reporting areas Number`
* `Continuously reporting areas Number`
* `Continuously reporting areas Rate`
* `Continuously reporting areas Ratio`

```{r, message = FALSE, warning = FALSE}
# Rename all the variables in AnnualReport_Raw
names(AnnualReport_Raw)[2] <- paste("Selected reporting areas Number")
names(AnnualReport_Raw)[3] <- paste("Continuously reporting areas Number")
names(AnnualReport_Raw)[4] <- paste("Continuously reporting areas Rate")
names(AnnualReport_Raw)[5] <- paste("Continuously reporting areas Ratio")

# Remove the redundant row of column names, create regular expression to grasp all the numerical cell values and convert those values from strings to numbers
AnnualReport_Clean <-
  AnnualReport_Raw %>%
  filter(row_number() > 1) %>%
  mutate(`Selected reporting areas Number` = gsub(pattern = ",", replacement = "", x = `Selected reporting areas Number`)) %>%
  extract(col = `Selected reporting areas Number`, into = "Selected reporting areas Number", regex ="(^\\d+)") %>%
  mutate(`Selected reporting areas Number` = as.numeric(`Selected reporting areas Number`)) %>%
  mutate(`Continuously reporting areas Number` = gsub(pattern = ",", replacement = "", x = `Continuously reporting areas Number`)) %>%
  extract(col = `Continuously reporting areas Number`, into = "Continuously reporting areas Number", regex ="(^\\d+)") %>%
  mutate(`Continuously reporting areas Number` = as.numeric(`Continuously reporting areas Number`)) %>%
  mutate(`Continuously reporting areas Rate` = as.numeric(`Continuously reporting areas Rate`)) %>%
  mutate(`Continuously reporting areas Ratio` = as.numeric(`Continuously reporting areas Ratio`))

str(AnnualReport_Clean)
```


# Data Visualization and Analysis

#### 1. Map for Trap laws distribution across USA in 2016

**Variables**

* `states`
* `Anti-Choice Law`
```{r, message = FALSE, warning = FALSE, fig.height = 20, fig.width = 12, fig.align = 'center'}
# Create a map indicating Trap Laws distribution across USA 
TrapLaws_Map <-
  plot_usmap(region = 'states', data = TrapLaw, values = "Anit-Choice Law", labels = TRUE) +
  labs(title = "Trap Laws in USA", subtitle = "A distribution of Trap Laws across USA in 2016", fill = "Trap Law") +
  scale_fill_manual(values = c("#FF7F50", "#5ab4ac")) +
  theme(legend.position = "right", legend.background = element_rect(fill = "#e6e6ff", size = 0.5, linetype = "solid"))
TrapLaws_Map
# Save the map plot
ggsave("TrapLaws_Map.png", height = 15, width = 12)
```
**Analysis**

* **Trap Laws have been popularized across USA since 2007 and reached peak in 2016(Trump’s administration tried to push back Roe v. Wade)**
* **Up to 2016, There are only 6 remaining States that do not have anti-choice laws against abortion: Oregon, Montana, Colorado, West Virginia, Vermont, New Hampshire**

#### 2. Annual Report

**Variables**

* `Year`
* `Selected reporting areas Number`
* `Continuously reporting areas Number`
```{r, message = FALSE, message = FALSE}
AnnualReport_Clean %>%
  select(`Selected reporting areas Number`) %>%
  summarise(total = sum(`Selected reporting areas Number`), max_selected = max(`Selected reporting areas Number`), selected = min(`Selected reporting areas Number`), mean_selected = mean(`Selected reporting areas Number`))

AnnualReport_Clean %>%
  select(`Continuously reporting areas Number`) %>%
  summarise(total = sum(`Continuously reporting areas Number`), max_continuoulsy = max(`Continuously reporting areas Number`), min_continuously = min(`Continuously reporting areas Number`), mean_continuously = mean(`Continuously reporting areas Number`))
```

```{r, message = FALSE, warning = FALSE, fig.height = 8, fig.width = 10, fig.align = 'center'}
AnnualReport_Clean %>%
  ggplot(aes(x = Year)) +
  geom_line(aes(y = `Selected reporting areas Number`, group = 1, color = "red")) +
  geom_point(aes(y = `Selected reporting areas Number`, group = 1, color = "red")) +
  geom_line(aes(y = `Continuously reporting areas Number`, group = 1, color = "blue")) +
  geom_point(aes(y = `Continuously reporting areas Number`, group = 1, color = "blue")) +
  labs(title = "Reported Legal Abortion 2007-2016", col = "Areas") +
  scale_color_manual(labels = c("Selected reporting areas Number", "Continuously reporting areas Number"), values = c("red", "blue")) +
  scale_y_continuous(name = "Selected reporting areas", sec.axis = sec_axis((~ . + 10), name = "Continuously reporting areas")) +
  theme(legend.position = "right", legend.background = element_rect(fill = "#e6e6ff", size = 0.5, linetype = "solid")) +

ggsave("AnnualReport.png", height = 8, width = 10)
```

#### 3. States Report Comparison

**Variables**

* `State/Area`
* `Residence Number`
* `Occurrence Number`
```{r, message = FALSE, message = FALSE}
StatesReport2015_Clean %>%
  select(`Residence Number`, `Occurrence Number`) %>%
  na.omit() %>%
  summarise(max_residence = max(`Residence Number`), min_residence = min(`Residence Number`), mean_residence = mean(`Residence Number`), max_occurrence = max(`Occurrence Number`), min_occurrence = min(`Occurrence Number`), mean_occurence = mean(`Occurrence Number`))

StatesReport2016_Clean %>%
  select(`Residence Number`, `Occurrence Number`) %>%
  na.omit() %>%
  summarise(max_residence = max(`Residence Number`), min_residence = min(`Residence Number`), mean_residence = mean(`Residence Number`), max_occurrence = max(`Occurrence Number`), min_occurrence = min(`Occurrence Number`), mean_occurence = mean(`Occurrence Number`))
```

```{r, message = FALSE, warming = FALSE, fig.height = 8, fig.width = 16, fig.align = 'center'}
StatesReport2015_Clean_Year <-
  StatesReport2015_Clean %>%
  mutate(Year = rep(x = "2015", times = 53))

StatesReport2016_Clean_Year <-
  StatesReport2016_Clean %>%
  mutate(Year = rep(x ="2016", times = 53))

StatesReport_Comparison <-
  rbind(StatesReport2015_Clean_Year, StatesReport2016_Clean_Year) %>%
  pivot_longer(cols = c(`Residence Number`, `Occurrence Number`), names_to = "State Report", values_to = "Count")

StatesReport_Comparison %>%
  na.omit() %>%
  ggplot(aes(x = `State/Area`, y = Count)) +
  geom_bar(stat = "identity", position = 'dodge', aes(color = `State Report`, fill = `State Report`, alpha = 0.25)) +
  labs(title = "States Report Comparison 2015 & 2016") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), legend.position = "bottom", legend.background = element_rect(fill = "#e6e6ff", size = 0.5, linetype = "solid")) +
  facet_grid(Year ~.)

ggsave("StatesReportComparison.png", height = 8, width = 16)
```

#### 4. 2016 States Report Dendrogram

**Variables**

* `State/Area`
* `Residence Number`
* `Residence Rate`
* `Residence Ratio`
* `Occurrence Number`
* `Occurrence Rate`
* `Occurrence Ratio`
* `% obtained by out-of-state residents`
```{r, message = FALSE, warming = FALSE, fig.height = 6, fig.width = 8, fig.align = 'center'}
StatesReport2016_Clean2 <-
  na.omit(StatesReport2016_Clean)

StatesReport2016_Clean2$`State/Area` = as.numeric(as.factor(StatesReport2016_Clean2$`State/Area`))

row.names(StatesReport2016_Clean2) <- c("Alabama", "Alaska", "Arizona", "Arkansas", "Colorado", "Connecticut", "Delaware", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin")

Distance <-
  dist(scale(StatesReport2016_Clean2), method = "euclidean")

# Create a vector named highlight which contains 5 No Trap Law States
highlight <- c("Oregon", "Montana", "Colorado", "West Virginia", "Vermont" )

# Define a function named ColorLeaf to highlight specific labels in Dendrogram
ColorLeaf <- function(x) {
  if (is.leaf(x) && attr(x, "label") %in% highlight) {
    attr(x, "nodePar") <- list(lab.col="#FF7F50", pch=NA)
  }
  return(x)
}

hc <- hclust(Distance, method = "ward.D2")

Dendrogram <- dendrapply(as.dendrogram(hc), ColorLeaf)

plot(Dendrogram)
```

#### 5. No Trap Law States Report Comparison

**Variables**

* `State/Area`
* `Residence Number`
* `Occurrence Number`
```{r, message = FALSE, warming = FALSE, fig.height = 6, fig.width = 8, fig.align = 'center'}
NoTrapLaw_StatesReport_2015 <-
  StatesReport2015_Clean %>%
  filter(`State/Area` %in% c("Oregon", "Montana", "Colorado", "West Virginia", "Vermont", "New Hampshire¶")) %>%
  mutate(Year = rep(x = "2015", times = 6))

NoTrapLaw_StatesReport_2016 <-
  StatesReport2016_Clean %>%
  filter(`State/Area` %in% c("Oregon", "Montana", "Colorado", "West Virginia", "Vermont", "New Hampshire¶")) %>%
  mutate(Year = rep(x = "2016", times = 6))

NoTrapLawStates_Comparison <-
  rbind(NoTrapLaw_StatesReport_2015, NoTrapLaw_StatesReport_2016) %>%
  pivot_longer(cols = c(`Residence Number`, `Occurrence Number`), names_to = "State Report", values_to = "Count")

NoTrapLawStates_Comparison %>%
  na.omit() %>%
  ggplot(aes(x = `State/Area`, y = Count)) +
  geom_bar(stat = "identity", position = 'dodge', aes(color = `State Report`, fill = `State Report`, alpha = 0.25)) +
  labs(title = "No Trap Law States Comparison 2015 & 2016") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), legend.position = "bottom", legend.background = element_rect(fill = "#e6e6ff", size = 0.5, linetype = "solid")) +
  facet_grid(Year ~.)

ggsave("NoTrapLawStatesReportComparison.png", height = 6, width = 8)
```


#### 6. Age Group Comparison in USA

**Variables**

* `State/Area`
* `Year`
* `Age Group`
* `Count`
```{r, message = FALSE, warming = FALSE, fig.height = 8, fig.width = 10, fig.align = 'center'}
AgeGroup2015_Clean_Year <-
  AgeGroup2015_Clean %>%
  mutate(Year = rep(x = "2015", times = 48))

AgeGroup2016_Clean_Year <-
  AgeGroup2016_Clean %>%
  mutate(Year = rep(x = "2016", times = 47))

AgeGroup_Comparison_USA <-
  rbind(AgeGroup2015_Clean_Year, AgeGroup2016_Clean_Year) %>%
  pivot_longer(cols = c(`Age group yrs (<15)`, `Age group yrs (15–19)`, `Age group yrs (20–24)`, `Age group yrs (25–29)`, `Age group yrs (30–34)`, `Age group yrs (35–39)`, `Age group yrs (>=40)`), names_to ="Age Group", values_to = "Count")

AgeGroup_Comparison_USA %>%
  na.omit() %>%
  ggplot(aes(x = `Age Group`, y = Count)) +
  geom_boxplot(aes(color = `Age Group`, fill = `Age Group`), notch = TRUE, alpha = 0.25, outlier.size = 2, outlier.colour = "gray") + 
  facet_wrap( ~ Year ) +
  ylim(0, 12000) +
  labs(title = "Age Group Comparison 2015 & 2016 in USA") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), legend.position = "bottom", legend.background = element_rect(fill = "#e6e6ff", size = 0.5, linetype = "solid"))

ggsave("AgeGroupComparison_USA.png", height = 8, width = 10)
```

#### 7. Age Group Comparison in No Trap Law States

**Variables**

* `State/Area`
* `Total abortions reported by known age`
* `Year`
* `Age Group`
* `Count`
```{r, message = FALSE, warming = FALSE, fig.height = 6, fig.width = 8, fig.align = 'center'}
NoTrapLaw_AgeGroup2015 <-
  AgeGroup2015_Clean %>%
  filter(`State/Area` %in% c("Oregon", "Montana", "Colorado", "West Virginia", "Vermont")) %>%
  mutate(Year = rep(x = "2015", times = 5))

NoTrapLaw_AgeGroup_2016 <-
  AgeGroup2016_Clean %>%
  filter(`State/Area` %in% c("Oregon", "Montana", "Colorado", "West Virginia", "Vermont")) %>%
  mutate(Year = rep(x = "2016", times = 5))
  

NoTrapLaw_AgeGroup_Comparison <-
  rbind(NoTrapLaw_AgeGroup2015, NoTrapLaw_AgeGroup_2016) %>%
  pivot_longer(cols = c(`Age group yrs (<15)`, `Age group yrs (15–19)`, `Age group yrs (20–24)`, `Age group yrs (25–29)`, `Age group yrs (30–34)`, `Age group yrs (35–39)`, `Age group yrs (>=40)`), names_to ="Age Group", values_to = "Count")

NoTrapLaw_AgeGroup_Comparison %>%
  na.omit() %>%
  ggplot(aes(x = `Age Group`, y = Count)) +
  geom_bar(stat = "identity", position = 'stack', aes(color = `Age Group`, fill = `Age Group`, alpha = 0.25)) +
  labs(title = "Age Group Comparison 2015 & 2016 in No Trap Law States") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1), legend.position = "bottom", legend.background = element_rect(fill = "#e6e6ff", size = 0.5, linetype = "solid")) +
  facet_wrap( ~ Year )

ggsave("NoTrapLawAgeGroupComparison.png", height = 6, width = 8)
```














