---
title: "STAT 184 Final Project"
author: "Yiwen Huang"
date: "Sunday by 11:59pm"
output: html_notebook
---


# Set Up
```{r, message = FALSE, warning = FALSE}
# clean up workspace environment
rm(list = ls())

# all packages used for the assignment
library(DataComputing)
library(mosaic)
library(tidyverse)
library(lubridate)
library(readxl)
library(rvest)
library(knitr)
library(mosaic)
library(ggplot2)
library(usmap)
library(rpart.plot)
library(rpart)
```


# Research Question

### Are Trap Laws associated with lower reported legal abortion number?

* Trap Law: Targeted restrictions on abortion providers (TRAP) laws are costly, severe, and medically unnecessary requirements imposed on abortion providers and woman's health centers. Often pushed by anti-abortion politicians under the guise of "woman's health," the real aim of TRAP laws is to shut down abortion providers and make it more difficult for people to access abortion.


# Data Access

### Import Data Source

#### 1. Abortion Surveillance — United States, 2015
```{r, message = FALSE, warning = FALSE}
# Assign the url for source 1 to variable url1
url1 <- "https://www.cdc.gov/mmwr/volumes/67/ss/ss6713a1.htm?fbclid=IwAR1RVZmh6lh4UWS6z6jL_fwbVm0pfCJ9H9azAK8TLS5hwyss-8qMi5uG9vo"

# Scrape all HTML tables found on the web page into a "list" object named Abortion2015 in the R environment
Abortion2015 <- url1 %>%
  read_html(header = TRUE) %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)
```

#### 2. Abortion Surveillance — United States, 2016
```{r, message = FALSE, warning = FALSE}
# Assign the url for source 2 to variable url2
url2 <- "https://www.cdc.gov/mmwr/volumes/68/ss/ss6811a1.htm#T1_down"

# Scrape all HTML tables found on the web page into a "list" object named Abortion2016 in the R environment
Abortion2016 <- url2 %>%
  read_html(header = TRUE) %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)
```

#### 3. Anti_Choice Law by States
```{r, message = FALSE, warning = FALSE}
# Use function read-excel from package readxl to read a local excel file into an R object named TrapLaw
TrapLaw <- read_excel("/Users/yiwen/Desktop/FALL 2020/STAT 184/Final Project/TrapLaw.xlsx")
```























